A simple probabilistic approach to finding strongly lensed quasars in ground-based imaging

### 2013-01-25 docs/paper1.tex

\title[Automated Gravitational Lens Detection]
{Automated Detection of Gravitationally Lensed Quasars in Multi-epoch
Ground-based Imaging Data}

@@

\begin{abstract}
\end{abstract}

>>>>

\title[finding lensed quasars]{A simple probabilistic approach to finding strongly lensed quasars in ground-based imaging}

@@

\begin{abstract}
In some heuristic sense, we end up believing that a particular source
is a galaxy-scale strong gravitational lens because it is more
plausible to explain it as a superposition of a gravitationally
distorted background source and a foreground lensing galaxy than it is
to explain it as a single object with very odd morphology.  Here we
construct a quantitative approximation to this comparison of
plausibilities by fitting multi-epoch, multi-band ground-based imaging
of gravitational-lens candidates with two image models, one of which
is a strongly lensed point source and simple lensing galaxy, and the
other is a complex galaxy constructed from a simple galaxy and $K$
point sources.  The relative plausibility of the lensing hypothesis is
approximated by a likelihood ratio between these models, evaluated at
best-fit (that is, not marginalized), penalized by a factor accounting
for relative model freedom.  We fit these models and compute lensing
plausibility for four candidate lens sources with \panstarrs\ $grizY$
imaging.  Two of the sources are known lenses, and two are not; we
find we can confidently identify the lenses and rule out the
non-lenses with our plausibility.  Even though the plausibility
calculation involves fitting highly parametrized models of
astronomical sources and the varying point-spread function in many
pixels of imaging, it can be calculated in seconds on contemporary
hardware; it can be run at scale on lens candidates in \panstarrs\ and
other surveys.
\end{abstract}
